### 2017-03-01 doc/Thesis/data/theory.tex
some last minute update to the theory section

Where $n$ is the time step, $k$ is the iteration number, and $\epsilon_{\varphi}$ is the numerical criteria provided by a user. Through testing, iterations do not always converge this criteria; in this circumstance, max number of iterations is enforced.

@@

An extension of implicit discretization is the backward difference formulae (BDF) \cite{Gear:2007}.  BDF's can increase the order of error convergence by interpolating solutions from previous time time steps. The general formula for BDF is described by \eqt{eq:bdf}.
\be 
\varphi_{n+1} = (IV-\Delta t A_{n+1})^{-1}\left[IV\sum_{j=0}^{k-1}\alpha_{jk}\varphi_{n-j} +
\Delta t b_{n+1}\right]
\label{eq:bdf}
\ee
Where $k$ is the order of the error convergence and $\alpha_{ij}$ are coefficients chosen such that the temporal truncation error is minimized.  The benefit of using BDF is that any order of convergence can be applied and are all unconditionally stable, so the IQS approximation can easily be validated for high order discretization.S

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Singly-Diagonally-Implicit Runge-Kutta Method}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Linear Solution Techniques}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

MOOSE contains a suite of solution techniques to evaluate a linear system generated by the Rattlesnake action system. For most neutronics and reactor applications, direct evaluation of the system is overbearingly expensive, both computationally and by memory usage. Therefore, the most common technique (default in MOOSE) is PJFNK. This technique is briefly described in \sct{sect:pjfnk}. PJFNK entails approximately solving the Jacobian action on the solution with a finite difference approach using GMRES iteration. Each GMRES iteration requires a residual evaluation, which becomes globally expensive with many iterations. Employing a preconditioner on the solution can decrease the number of iterations and save execution time. The purpose of the preconditioner is to resemble the Jacobian (reduce number of iterations), while being easily invertable (preconditioner overhead). MOOSE has four useful assembly methods for building an effective preconditioner:

@@

When developing a kernel in MOOSE, a function is implement that evaluates the Jacobian for each variable it depends on. In a multi-variable system, like multigroup diffusion, the Jacobian contains "blocks" for each variable and the coupling between them. Diagonal blocks contain the variable contribution to itself, i.e. diffusion and removal. The off-diagonal blocks contain variable-to-variable contributions, i.e. scattering and fission. The default method assembles only the diagonal blocks; this preconditioner is relatively easy to invert, but is not optimal for reducing linear iterations. SMP assembles the diagonal blocks and any off-diagonal blocks specified by the user. Assembling more off-diagonal blocks reduces linear iterations (inputting all blocks requires only one GMRES iteration for a linear system), but increases the expense of inverting the preconditioner. Therefore, SMP requires extensive analysis for optimization. FDP simply does a finite difference approximation for the Jacobian based on residual evaluations. This method only requires one GMRES iteration for a linear system, but is impractically expensive. PBP is an advanced adaptation to SMP. This method assembles the preconditioner similarly to SMP, but each block is inverted as a independent system and iteratively communicates with the rest of the blocks. PBP can be effective for a lower block diagonal system, i.e. including only downscattering blocks, because no communication iterations are required. Selecting an effective preconditioner and assembly method is essential for maximizing computational efficiency of large reactor simulations.

>>>>

Where $n$ is the time step, $k$ is the iteration number, and $\epsilon_{\varphi}$ is the numerical criteria provided by a user. Through testing, iterations do not always converge this criteria; in this circumstance, the maximum number of iterations is reached.

@@

An extension of implicit discretization is the backward difference formulae (BDF) \cite{Gear:2007}.  BDF's can increase the order of error convergence by interpolating solutions from previous time time steps. The general formula for BDF is described by \eqt{eq:bdf}.
\be 
\varphi_{n+1} = (IV-\Delta t A_{n+1})^{-1}\left[IV\sum_{j=0}^{k-1}\alpha_{j}^{k}\varphi_{n-j} +
\Delta t \alpha_{k+1}^{k} b_{n+1}\right]
\label{eq:bdf}
\ee
where $k$ is the order of the error convergence and $\alpha_{j}^{k}$ (vector of length $k+1$) are coefficients chosen such that the temporal truncation error is minimized.  The benefit of using BDF is that any order of convergence can be applied and are all unconditionally stable, so the IQS approximation can easily be validated for high order discretization.
\tcr{I think one term was missing in your eq 2.35. also, it would be good to given the alpha values for the methods you implemented}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Singly-Diagonally-Implicit Runge-Kutta Method}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Linear Solution Techniques}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\tcr{JFNK is really for nonlinear systems. I would avoid confusing the reader here. I would start with PJFNK and Picard are
the nonlinear solvers in MOOSE. Each of the nonlinear solvers requires linear solves ... and then you talk about linear solves ...}

MOOSE contains a suite of solution techniques to solve the linear system generated by the Rattlesnake action system. For most neutronics and reactor applications, direct inversion (e.g., LU solve) of the system is overbearingly expensive, both computationally and by memory usage. Therefore, the most common technique (default in MOOSE) is PJFNK. This technique is briefly described in \sct{sect:pjfnk}. PJFNK entails approximately solving the Jacobian action on the solution with a finite difference approach using GMRES iteration. Each GMRES iteration requires a residual evaluation, which becomes globally expensive with many iterations. Employing a preconditioner on the solution can decrease the number of iterations and save execution time. The purpose of the preconditioner is to resemble the Jacobian (reduce number of iterations), while being easily invertable (preconditioner overhead). MOOSE has four useful assembly methods for building an effective preconditioner:

@@

When developing a kernel in MOOSE, a function is implement that evaluates the Jacobian for each variable it depends on. In a multi-variable system, like multigroup diffusion, the Jacobian contains "blocks" for each variable and the coupling between them. Diagonal blocks contain the variable contribution to itself, i.e., diffusion and removal. The off-diagonal blocks contain variable-to-variable contributions, i.e. scattering and fission. The default method assembles only the diagonal blocks; this preconditioner is relatively easy to invert, but is not optimal for reducing linear iterations. SMP assembles the diagonal blocks and any off-diagonal blocks specified by the user. Assembling more off-diagonal blocks reduces linear iterations (inputting all blocks requires only one GMRES iteration for a linear system), but increases the expense of inverting the preconditioner. Therefore, SMP requires extensive analysis for optimization. FDP simply does a finite difference approximation for the Jacobian based on residual evaluations. This method only requires one GMRES iteration for a linear system, but is impractically expensive. PBP is an advanced adaptation to SMP. This method assembles the preconditioner similarly to SMP, but each block is inverted as a independent system and iteratively communicates with the rest of the blocks. PBP can be effective for a lower block diagonal system, i.e., including only downscattering blocks, because no communication iterations are required. Selecting an effective preconditioner and assembly method is essential for maximizing computational efficiency of large reactor simulations.
